Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities

### 2020-09-16 writing/transect_paper.Rmd
address most reviewer comments in intro

@@

Fungi are key drivers of nutrient cycling in terrestrial ecosystems.
One important guild of fungi form ectomycorrhizas (ECM), a symbiosis between fungi and plants in which fungal hyphae enclose the plant's fine root tips.
The fungi provide nutrients and protection from pathogens in exchange for carbon from the plant [@Smith2010].
Approximately 8% of described fungal species are thought to take part in ECM symbiosis [@ainsworth2008ainsworth;@Rinaldi2008].
Although only about 2% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate and boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, *Uapaca* (Phyllanthaceae) and Fabaceae tr. Amherstieae [@brundrett2017a].

@@

As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
Fungal metabarcoding studies using short-read HTS technologies have targeted the ITS1 or ITS2 regions, which provide sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis [@Lindahl2013].
If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST [@altschul1990;@Lindahl2013].
However, this approach comes with some potential weaknesses.

While ITS1 and ITS2 often have suitable variation to distinguish species, they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa.
Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi [@tedersoo2015].

Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
In contrast, more recent denoising methods such as DADA2 [@callahan2017], Deblur [@amir2017], and UNOISE2 [@edgar2016] utilize read quality information to control for sequencing error while preserving intra-species variation.
The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and are less dataset specific [@callahan2017].

Assignment of taxonomic identities using BLAST requires *a priori* choice of thresholds for different taxonomic ranks.
Several algorithms specifically designed for taxonomic assignment have been published which use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments [e.g.,@wang2007;@edgar2016a;@murali2018a].
In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments [@somervuo2016;@gdanetz2017].

Recent long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions [@Tedersoo2018].
This can potentially improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis.
Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
Because OTU clustering may both "clump" different species into a single OTU, and "split" a single species into multiple OTUs [@ryberg2015], diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
In contrast, phylogenetic community distance measures [@wong2016] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree [@matsen2010; @berger2011], but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018].
However, long amplicon sequencing allows the inclusion of alignable regions for construction of more fully resolved phylogenetic trees directly from metabarcoding reads.
However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].

Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated Soudanian woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA) and reference databases (Unite, Warcup, RDP), and two different community distance measures (Bray-Curtis vs. weighted UNIFRAC).
We also present new algorithms for dividing the rDNA into regions, combining denoising results from multiple regions, and incorporating phylogenetic information into taxonomic assignments.

@@

Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
Both sites were located in woodlands dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
At each site, 25\ soil samples were collected along a linear transect at intervals of 1\ m in May 2015.
One third of the sample locations (3\ m spacing) were resampled one year later in June 2016.
For each sample, any coarse organic debris <!--Do you have any definition? Wider than 5 mm? MR-->was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade.
Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
Approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of lysis buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA) and lysed in the field using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
<!--The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis. -->

An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
Sequencing results for these samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.

@@

The gITS7 primers for the short amplicon were indexed for multiplexing (Supplementary File 1).
Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer, 150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
The reaction conditions were 10\ min at 95°, followed by 35\ cycles of 60\ s at 95°, 45\ s at 56°, and 50\ s at 72°, and finally 3\ min at 72°.
Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity, which were pooled after amplification.

@@

Amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752\ 37 UPPSALA, Sweden).
Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.

Additionally, the short amplicon libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300\ bp at the SNP&SEQ Technology Platform (Dept. of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751\ 44 UPPSALA, Sweden).
The Illumina library was pooled with samples for another project, with half of the reads from one lane devoted to the current study.

@@

We attempted to denoise both long and short PacBio amplicons using DADA2 according to the steps outlined in the supplementary information in @callahan2019.
However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = ",")`.
Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r format(readd(demuxcounts)["pb_500"], format = "d", big.mark = ",")` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:

Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
First, the 5.8S rDNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013], and all bases before the 5.8S were assigned to ITS1.
No attempt was made to remove the approximately 12\ bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
<!--This approach was able to identify both the 5' and 3' ends of the 5.8S rDNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
A CM was generated from the alignment using `cmbuild` from Infernal.
The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using `cmalign` from Infernal.
The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in Figure\ \@ref(fig:rDNA),
where LSU1-4 represent the conserved regions of LSU flanking the variable D1-3 regions [@michot1984].
For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
Code to extract the regions, including annotated seed alignments and CMs, is available in the new R package `LSUx`.

>>>>

<!-- R1: This study seeks to determine how both sequencing target length and technology affect the spatial and temporal turnover of soil fungal communities in an African forest. The authors take an ambitious approach to these research questions by comparing multiple short-read technologies (Illumina, PacBio, and Ion Torrent) to a related long-read technology (PacBio), allowing them to make many valuable comparisons across datasets. Along with field sample replication, both spatially and temporally, the authors carefully consider a number of different bioinformatic steps for each sample, e.g. among taxonomic databases as well as whether to cluster ASV data or not. Importantly, after discovering that long read data has inherent properties that make it poorly suited for the popular DADA2 bioinformatic pipeline, the authors generate a series of new R-based tools to analyze long read data in ways that greater increase data retention. Further, they use a clever set of covariance models to improve the taxonomic assignments of ASVs that are not assigned at the bulk taxonomy step. This additional step is important in increasing the amount of sequence data that can be annotated to putative functional guilds, which in this case focuses on ECM fungi.

Overall, I found the study to be a really excellent demonstration of how different sequencing technologies and target lengths can be compared for fungi. The overall conclusion that the ecological outcome is largely similar between the different datasets is heartening. I am quite sure that longer reads will continue to increase in popularity as costs decline and this study represents an important bridge in that transition – both for its careful benchmarking but also for its presentation of tools that allow for better incorporation of long read sequence data into existing bioinformatic pipelines. I hope that my comments below help the authors make some relatively minor improvements, particular in adding more biological interpretation in a few places to help more deeply understand the patterns they have identified. -->

<!-- R2: The authors have collected an interesting data set but I feel strongly that the work requires significant improvement to the Introduction (particularly the justification) and Discussion (particularly some explanations, even speculative) sections, as detailed below in my specific comments.

Regarding the justification and conclusions, a quick scan of the literature revealed a study that used long(er) reads (covering SSU, ITS, LSU; Heeger et al. 2019 in this journal) and also separated these prior to clustering and assignment. I therefore see the contribution as stated here (line 614) as overstated except that denoising was applied prior to clustering. Phylogenetic reconstruction was also used here, although the conclusions are not justified (see my comment below). I see an important contribution of this work as developing a new method to separate fragments; however, although it should be clear (i.e. examined in the Introduction) why existing approaches like ITSx or others are not suitable.

"Long-read" has become a vague descriptor and should be clearly defined early on for this manuscript (in the abstract). How many bp or markers? I have seen the term used in metabarcoding to mean anything from a few bp longer than a MiSeq PE 2x250, to nearly 5000 bp covering multiple rRNA genes.-->

<!-- R3: This paper analyzes a transect of soils in Africa and reports how results differ from 3 differing sequencing technologies and variations to bioinformatics platforms. With sequencing technologies changing so rapidly, and so many methods to choose from, I’m sure this paper would be of interest to the Molecular Ecology Resources audience, as it provides useful data comparing sequencing technologies. I congratulate the author on herculean efforts to write so many R packages, for making his packages and scripts publicly available, and for making a pipeline that allegedly improves ASV recovery for long amplicons. I personally would find the paper more interesting if a treatment other than spatial turnover had been applied to these samples, like before and after a disturbance, or a different host tree, and one could examine if the treatment of interest gave similar inferences with different sequencing technologies. However, I believe their claim that sequencing technologies yield similar results. I do think the paper would be improved by adding in some alpha diversity analyses, which are likely to be less robust to vagaries of sequencing technologies than beta diversity analyses. I also did not see any discussion of normalization for variation in number of sequences per sample, which needs to be accounted for. Instead of figure 2 (which can be moved to the supplements), I would prefer instead to see a figure about correlations of alpha diversity among methods. I also think species accumulation curves would be useful to see. I have also suggested several ways the writing could be slightly altered to make it clearer and more engaging for the reader. The results section is quite dry, reporting the statistics rather than reporting the interesting results and using the statistics to support it. For example, “Microbial communities are spatially autocorrelated (R2 = 0.90, p <0.05)” is more interesting to read than the “Bray-curtis dissimilarity showed…” If some of the results section could be re-written to focus on the results rather than reporting statistics, it would be more engaging. I also think the introduction needs to be edited to be more cohesive with the rest of the document. For example, only abbreviations were used in the introduction but they were fully spelled out at numerous points in the methods, and this made the introduction unnecessarily difficult to follow. I also found that the discussion repeated the results a lot and thus could be truncated or re-focused to set the results more in the context of the current literature (for example some citations of similar papers was missing). Finally, I will freely admit that some portions of the methods section, like the discussion of LSUx, went over my head and were very difficult for me to understand and/or follow. I’m not sure if someone more bioinformatically inclined could do a better job reviewing some of the more technical portions or if they just need to be written more clearly. I actually spent 3x the amount of time I usually devote to reviews on this paper, and I made a lot of efforts to provide constructive criticisms of how this paper could be improved to be clearer for the reader, so I hope that they are helpful. See specific line comments below.-->

@@

Fungi are key drivers of nutrient cycling in terrestrial ecosystems.
One important guild of fungi form ectomycorrhizas (ECM), a symbiosis between fungi and plants in which fungal hyphae enclose the plant's fine root tips.
The fungi provide nutrients and protection from pathogens in exchange for carbon from the plant [@Smith2010].
Approximately 8% of described fungal species are thought to take part in ECM symbiosis [@ainsworth2008ainsworth;@Rinaldi2008].
Although only about 2% of land plant species form ECM, these include ecologically and economically important stand-forming trees belonging to both temperate and boreal groups such as Pinaceae and Fagaceae, and tropical groups such as Dipterocarpaceae, *Uapaca* (Phyllanthaceae) and Fabaceae tr. Amherstieae [@brundrett2017a], which together represent approximately 60% of tree stems globally [@steidinger2019].

@@

As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
Fungal metabarcoding studies using short-read HTS technologies such as 454, Illumina, and Ion Torrent
have usually targeted the rDNA internal transcribed spacer regions ITS1 or ITS2,
which are the standard molecular barcode for fungi, providing sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS [@schoch2012;@Lindahl2013].
In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice [@opik2010], and variable regions of the rDNA large subunit (LSU) have also been used for barcoding [@house2016;@tedersoo2015].
The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis [@Lindahl2013].
If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST [@altschul1990;@Lindahl2013] with public databases such as Genbank [@benson2013] and Unite [@nilsson2019a].
However, this approach comes with some potential weaknesses, discussed below.

ITS1 and ITS2 often have suitable variation to distinguish species, although closely related species may share identical ITS sequences in certain groups such as various Pezizomycotina [@schoch2012],
but this variablity means that they cannot be reliably aligned over the fungal kingdom [@Lindahl2013;@Tedersoo2018].
<!-- R2: This is true, but the fungi are a large and diverse group and thus it is no surprise that a highly variable rRNA marker can not be aligned. Perhaps this manuscript could discuss how meaningful a phylogenetic approach to analysis of complete fungal communities is anyway. Perhaps building a tree for each phylum is possible (and a realistic evolutionary model could be used). I think that this could be a contribution here - to explore how this variation could be dealt with using some knowledge of the clade "fungi". (Partly addressed?) -->
Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa.
Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi [@tedersoo2015].

Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
In contrast, more recent denoising methods such as DADA2 [@callahan2017], Deblur [@amir2017], and UNOISE2 [@edgar2016] utilize read quality information to control for sequencing error while preserving intra-species variation.
The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, and have been suggested to be less dataset specific [@callahan2017].

Because both OTU clustering and denoised ASVs may "clump" different species into a single unit and "split" a single species into multiple units [@ryberg2015], diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
In contrast, phylogenetic community distance measures [@wong2016] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree [@matsen2010; @berger2011], but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments [@tedersoo2018], or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases [@fouquier2016].

Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
Using BLAST requires *a priori* choice of thresholds for different taxonomic ranks.
Although GenBank is often used as a first step for sequence identification, the database is ...
Mention Unite, RDP, Warcup.
<!-- R3: I would like to see a few more sentences of discussion of the issues with taxonomic assignments … I’m sure you know of them and are being brief for word limits…but I think it would be useful to mention more for young readers or people unfamiliar with fungi and issues you can come up against...like huge percentages of your sequences not getting names due to poor databases, multiple sequences getting the same names, multiple ITS copies within the same species getting different names, chimeric or incorrect taxonomic identifications in BLAST,  etc etc-->
Several algorithms specifically designed for taxonomic assignment have been published which use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments, including the RDP Classifier [RDPC, @wang2007], SINTAX [@edgar2016a] and IDTAXA [@murali2018a] among others.
In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments [@somervuo2016;@gdanetz2017].
<!-- R1: Palmer et al.(2018) use a similar ‘global’ approach to taxonomic assignment in the amptk pipeline that I think would be worth citing here. (They also discuss length variation in the ITS as a factor influencing taxa recovery, which is relevant to the points highlighted in the previous paragraph). -->

Recent long-read HTS technologies such as Pacific Biosciences Single Molecule Real Time sequencing (PacBio) enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions.
This can improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis [@Tedersoo2018].
Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number [@Kennedy2018].

Because of the variety of sequencing platforms and analytical pipelines which have been used in metabarcoding studies, comparisons between studies may be difficult.
Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
<!-- R2: This comes as a surprise - why use multiple sequencing strategies? There is no background given on the strategies used. Why use three different short-read platforms? Some discussion of how error rates differ and how the proprietary software (e.g., PacBio, Illumina) differs in its way of scoring is needed. The overall approach needs more justification in the Introduction section. I suggest either framing the question more in terms of using 1 vs 2 rRNA regions; although it must be placed in context based on the fact that other studies have used multiple rRNA regions in long-read studies); perhaps exploring in more detail how read depth affects the conclusions; however, read depth has not been mentioned in the Intro at all.-->
We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA)
and reference databases (Unite, Warcup, RDP),
and with non-phylogenetic and phylogenetic community distance measures.
We also present new algorithms for dividing the rDNA into regions, combining denoising results from multiple regions, and incorporating phylogenetic information into taxonomic assignments.
<!-- R2: The need for new algorithms is not introduced or established in the Introduction. For example, what are the limitation of ITSx?-->

@@

Sampling was conducted at two sites (Ang: N 9.75456° W 2.14064°; Gan: N 9.75678° W 2.31058 °) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
<!-- R3: I entered those locations into google maps. The first one took me to Ghana, the second took me to the Pacific Ocean. Something is off – you need to fix this.-->
Both sites were located in Soudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Caesalpinioideae).
At each site, 25\ soil samples were collected along a linear transect at intervals of 1\ m in May 2015.
<!-- R3: this is confusing. Do you mean that you had one 25m transect and 25 samples were taken along that one transect at 1m intervals?-->
One third of the sample locations (3\ m spacing) were resampled one year later in June 2016.
For each sample, any coarse organic debris <!--Do you have any definition? Wider than 5 mm? MR-->was removed from the soil surface and a sample of approximately 5cm×5cm×5cm was extracted with a sterilized knife blade.
Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
Approximately 250\ mg total of soil was collected from two locations in the homogenized soil sample and placed into a separate 2.0\ mL microtube containing 750\ mL of lysis buffer and lysis beads (Xpedition^TM^ Soil/Fecal DNA miniprep, Zymo Research Corporation, Irvine, California, USA) and lysed in the field using a handheld bead-beater (TerraLyser^TM^; Zymo Research Corporation).
<!--The remainder of each sample was dried at 40°C with an electric dehydrator within 24-48 hours of collection for bulk chemical analysis. -->
<!-- R3: why would you lyse in the field? Aren’t you concerned about needing to refrigerate or freeze your sample? This is highly unorthodox-->

An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
Sequencing results for these samples differed significantly (PERMANOVA with 9999 permutations, $p < `r readd(buffer_adonis)["buffer","p.value"] %>% formatC(format = "f")`$, $R^2 = `r readd(buffer_adonis)["buffer","R2"] %>% formatC(format = "f", digits = 2)`$) from samples preserved using the Xpedition^TM^ lysis buffer (Figures \@ref(fig:buffer-pcoa), \@ref(fig:buffer-compare-long), and \@ref(fig:buffer-compare-short)); as such these samples were excluded from our spatial analyses.
<!-- R3: which samples were excluded? The ones that were hand lysed or the ones that were preserved in lifeguard?-->
However, reads from these samples were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.

@@

The gITS7 primers for the short amplicon were indexed for multiplexing (Supplementary File 1).
Amplification was performed by polymerase chain reaction (PCR) in 20µl reactions containing 200\ µM dNTP mix, 250\ µM indexed gITS7 primer,
<!-- R3: so the gITS7 primer has a barcode and adaptor on it? I’m having a bit of trouble following your methods-->
150µM ITS4m, 2mM MgCl~2~, 0.1\ U *Taq* polymerase (Dream *Taq*, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7\ ng purified DNA in Dream *Taq* buffer.
The reaction conditions were 10\ min at 95°, followed by 35\ cycles of 60\ s at 95°, 45\ s at 56°, and 50\ s at 72°, and finally 3\ min at 72°.
Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity, which were pooled after amplification.

@@

Amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically as above.
<!-- R3: were technical replicates pooled prior to bead cleaning? Personally, I do not think technical replicates are necessary when using NGS (see Smith and Peay Plos One 2014) but I still need to be able to clearly follow your methods-->
An aliquot of 100\ ng of DNA from each sample (or the total PCR product if less than 100\ ng) was pooled into two libraries each for long and short amplicons.
Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences (PacBio) RS II sequencer at the Uppsala Genome Center (UGC; Uppsala Genome Center, Science
for Life Laboratory, Dept. of Immunology, Genetics and Pathology, Uppsala University, BMC, Box 815, SE-752\ 37 UPPSALA, Sweden).
Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.

Additionally, the short amplicon libraries were combined and sequenced using an Ion S5 (Ion Torrent) sequencer using one 520 chip at UGC, and a MiSeq (Illumina Inc.) sequencer using v3 chemistry with a paired-end read length of 300\ bp at the SNP&SEQ Technology Platform (Dept. of Medical Sciences, Uppsala University, BMC, Box 1432, SE-751\ 44 UPPSALA, Sweden).
The Illumina library was pooled with samples for another project, with half of the reads from one lane devoted to the current study.
<!-- R3: not sure that this line is necessary-->

@@

We attempted to denoise both long and short PacBio amplicons using DADA2 according to the steps outlined in the supplementary information in @callahan2019.
However, only `r readd(region_table)$ASVs[readd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(readd(region_table)$reads[readd(region_table)$region == "long"]/as.integer(gsub(" ", "", readd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
<!-- R3: Note: I have heard from many people that DADA2 performs poorly for ITS and that it was designed for 16S-->
We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = ",")`.
Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r format(readd(demuxcounts)["pb_500"], format = "d", big.mark = ",")` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.
We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:

<!-- R3: honestly this section is above my head and I can’t judge its accuracy. Hopefully another review is more bioinformatically inclined and can judge it better. -->
Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
First, the 5.8S rDNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013], and all bases before the 5.8S were assigned to ITS1.
No attempt was made to remove the approximately 12\ bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
<!--This approach was able to identify both the 5' and 3' ends of the 5.8S rDNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
<!-- R3: you spell out RDP and cite it here, but not in your first mention of it at line 109. I would switch this around.-->
A CM was generated from the alignment using `cmbuild` from Infernal.
The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using `cmalign` from Infernal.
The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved regions as shown in Figure\ \@ref(fig:rDNA),
where LSU1-4 represent the conserved regions of LSU flanking the variable D1-3 regions [@michot1984].
For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
Code to extract the regions, including annotated seed alignments and CMs, is available in the new R package `LSUx`.
<!-- R3: I tried to download LSUx but got an error. Can you do something about this?

> install.packages("LSUx")
And it threw the error
“Warning message:
package ‘LSUx’ is not available (for R version 4.0.2)”

is this the github repo? https://rdrr.io/github/brendanf/LSUx/src/R/lsux.R can you put a link to this instead or in addition?

If this is a methods paper, is the point to introduce LSUx? Can you provide a tutorial or better explanations of how one might actually implement LSUx? I have run analyses in QIIME2 and UPARSE many times but I don’t think I could figure out how to run your pipeline based on your paper here, and isn’t the point of this paper to make a new method available to your readers?-->
